%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Classicthesis Typographic Thesis
% LaTeX Template
% Version 1.3 (15/2/14)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% André Miede (http://www.miede.de)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
% General Tips:
% 1) Make sure to edit the classicthesis-config.file
% 2) New enumeration (A., B., C., etc in small caps): \begin{aenumerate} \end{aenumerate}
% 3) For margin notes: \marginpar or \graffito{}
% 4) Do not use bold fonts in this style, it is designed around them
% 5) Use tables as in the examples
% 6) See classicthesis-preamble.sty for useful commands
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[
		twoside,openright,titlepage,numbers=noenddot,headinclude,%1headlines,
                footinclude=true,cleardoublepage=empty,dottedtoc,
                BCOR=5mm,paper=a4,fontsize=11pt, % Binding correction, paper type and font size
                ngerman,american, % Languages
                ]{scrreprt} 
                
% Includes the file which contains all the document configurations and packages - make sure to edit this file
\input{config}

\begin{document}

\frenchspacing % Reduces space after periods to make text more compact

\raggedbottom % Makes all pages the height of the text on that page

\selectlanguage{american} % Select your default language - e.g. american or ngerman

%\renewcommand*{\bibname}{new name} % Uncomment to change the name of the bibliography
%\setbibpreamble{} % Uncomment to include a preamble to the bibliography - some text before the reference list starts

\pagenumbering{roman} % Roman page numbering prior to the start of the thesis content (i, ii, iii, etc)

\pagestyle{plain} % Suppress headers for the pre-content pages

%----------------------------------------------------------------------------------------
%	PRE-CONTENT THESIS PAGES
%----------------------------------------------------------------------------------------

\include{frontback/titlepage} % Main title page
\include{frontback/titleback} % Back of the title page
\cleardoublepage\include{frontback/opening} % Opening
\cleardoublepage\include{frontback/dedication} % Dedication page
\cleardoublepage\include{frontback/abstract} % Abstract page
\cleardoublepage\include{frontback/publication} % Publications from the thesis page
\cleardoublepage\include{frontback/acknowledgments} % Acknowledgements page

\pagestyle{scrheadings} % Show chapter titles as headings
\cleardoublepage\include{frontback/contents} % Contents, list of figures/tables/listings and acronyms

\cleardoublepage

\pagenumbering{arabic} % Arabic page numbering for thesis content (1, 2, 3, etc)
%\setcounter{page}{90} % Uncomment to manually start the page counter at an arbitrary value (for example if you wish to count the pre-content pages in the page count)

\cleardoublepage % Avoids problems with pdfbookmark

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - CHAPTERS
%----------------------------------------------------------------------------------------

\ctparttext{We begin this part with a general introduction that stresses the
    ever growing importance of cities in the world, and highlights the
    difficulties encountered when trying to reach a scientific understanding of
    these systems. We briefly outline the history of the quantitative tradition
    in the study of urban systems, and argue that we may be witnessing a second
quantitative revolution. We then  succintly present the methodology that we
followed during the past 3 years, and end this part with an outline of the
content presented in this thesis.} 

\part{Introduction} % First part of the thesis
\label{part:introduction}

\include{chapters/cities}
\include{chapters/quantitative-revolutions} % Chapter 1
\include{chapters/methodology}
\include{chapters/outline}

\cleardoublepage % Empty page before the start of the next part

%------------------------------------------------

\ctparttext{

The monocentric model of cities -- where all activities are organised around a
single activity center -- has pervaded the literature on urban systems for more
than $4$ decades. However, as it was repeatedly demonstrated, the model is empirically
inadequate.

The contribution of this part is threefold. First, we recount the history of
ideas about urban form, from the monocentric hypothesis and its origins, to the
various methods proposed to identify and count subcenters. We then demonstrate
empirically the existence of a polycentric transition for cities, and that the
number of centers increases as a sublinear function of population size. Finally,
we propose an out-of-equilibrium model that explains the emergence of new
subcenters as cities expand, and predicts the sublinear increase of the number
of centers with population size.

}

\part{Polycentri-city}
\label{part:polycentricity}

\include{chapters/monocentric}
\include{chapters/congestion}
\include{chapters/monocentric-discussion}
%------------------------------------------------

\ctparttext{The past decade has witnessed a renewed interest for the scaling of some of cities' characteristics with population size
    -- first discovered more than $60$ years ago.

The contribution of this part is threefold. First, we review the exisiting
literature on allometric scalings, sorting the measured
exponents by theme. We then propose a model to explain the scaling exponent of several
indicators related to mobility in cities, and discuss the theoretical and
practical consequences of these exponents.  Finally, we present some of the
challenges posed by scaling relationships: their interpretation,
and the issues they reveal about the definition of cities.}

\part{Scaling} % Second part of the thesis
\label{part:scaling}

\include{chapters/scaling-introduction} % Chapter 2
\include{chapters/mobility-to-scaling}
\include{chapters/scaling-implications}
%------------------------------------------------




\ctparttext{Residential segregation is a reality. A reality so rife that it has
    pervaded even our every day language though the expressions 'poor
    neighbourhood' or 'rich neighbouhood'. But despite its intuitive appeal,
    segregation is difficult to define.\\ 
    In this part, we propose to define segregation as a deviation to the
    unsegregated city, thereby providing a firm theoretical basis for any
    study of segregation patterns. We further propose
    a measure of attraction/repulsion of the different categories, which allows us to
    define unambiguously income classes from the original categories. We also study the
properties of neighbourhoods in which the different classes concentrate, and revisit the traditional poor
center/rich suburb dichotomy.} 

\part{Segregation}
\label{part:segregation}

\include{chapters/segregation-introduction}
\include{chapters/segregation}
%------------------------------------------------




\ctparttext{People, energy, information and goods are carried through cities
    (and across systems of cities) thanks to various networks. In this part, we
    succintly present our work on these---spatial---networks. 
    
    We first propose a quantative method to classify cities that is based on a new
    perspective on street patterns, and the use of the OpenStreetMaps database. In
    the second chapter, we propose a model for the growth of spatial networks
    based on cost-benefit analysis. The resulting networks exhibit a crossover
    between the star graph and the minimum spanning tree when the ratio of costs
    and benefits evolve. In the intermediate regime, the networks adopt a
    hub-and-spoke hierarchical structure that has many interesting properties.
    We conclude this part with a large-scale description of subway and
    railway networks. Using the model presented in the previous chapter, we are
    able to predict many of their properties based on the characteristics of the
underlying city or country.}

\part{Urban Networks} % Second part of the thesis
\label{part:networks}

\include{chapters/typology} % Chapter 2
\include{chapters/cost-benefit}
\include{chapters/scaling-networks}

%------------------------------------------------
%\ctparttext{Cities are not closed systems. Neither do they evolve in isolation.
%Rather, they interact in what has been called a system of cities. In this part,
%we focus on the movements of populations inside a system of city. We sutyd the
%network of residential migrations in the US, describe some of its properties,
%and show some unexpected consequences. Last, but not least, we show how Zipf's
%law finds a natural explanation in when considering the motion of people.} % Text on the Part 2 page describing the content in Part 2

%\part{Systems of cities} % Second part of the thesis

%\include{chapters/metropolitan-blend} % The metropolitan blend 
%\include{chapters/zipf} % Some work on Zipf
%\include{chapters/solving-mystery} % A model to explain Zipf's law

%------------------------------------------------


\ctparttext{\centering Self-explanatory title.} % Text on the Part 2 page describing the content in Part 2

\part{Conclusion} % Conclusion of the thesis
\label{part:conclusion}

\include{chapters/conclusion} % Only one chapter, that makes 13
%----------------------------------------------------------------------------------------
%	THESIS CONTENT - APPENDICES
%----------------------------------------------------------------------------------------

%\appendix

%\part{Appendix} % New part of the thesis for the appendix

%\include{Chapters/Chapter0A} % Appendix A
%\include{Chapters/Chapter0B} % Appendix B - empty template

%----------------------------------------------------------------------------------------
%	POST-CONTENT THESIS PAGES
%----------------------------------------------------------------------------------------

\cleardoublepage\include{frontback/bibliography} % Bibliography

%----------------------------------------------------------------------------------------

\end{document}
