% Introduction

\pdfbookmark[1]{Introduction}{Introduction}


\chapter{Quantitative revolution(s) in urban science}
\label{chap:quantitative_revolutions}

\begin{flushright}{\slshape    
And the first one now\\
Will later be last\\
For the times they are a-changin'} \\ \medskip
--- Bob Dylan 
\end{flushright}

\section{The first quantitative revolution}
\label{sec:the_first_quantitative_revolution}


\section{A second quantitative revolution?}
\label{sec:a_second_quantitative_revolution_}


    \subsection{New methods}
    \label{sub:new_methods}


    \subsection{New data}
    \label{sub:new_data}
    
    \subsection{A technological convergence}
    \label{sub:a_technological_convergence} 

It is difficult to make a concise summary of what is known and not known about
urban systems. The vast amount of knowledge that has been gathered so far seems
very little in comparison to the bewildering complexity of the object being
studied. Every map, every satellite view, every statistic, every step
in cities elicits a question yet to be answered. What do we have to answer them?
A surprisingly small array of empirical tools and models. A surprisingly small
amount of solid, undisputed empirical facts.\\

This pessimistic account is not to belittle the previous contributions to the
field, however. They did what is the most difficult: laying the conceptual
ground on which detailed enquiries can flourish. One thing I have learned, not
through reading philosophy of science books, or casual articles about the
scientific practice, but because I have been stuck myself, is that the most
difficult problems are conceptual ones. It is impossible to define a city
quantitatively a city before you have formed---with words, possibly drawings---a
conceptual picture of what a city is. It is impossible to study segregation
before you have logically clarified what one means by segregation---in the
spirit of analytic philosophy. However quantitative, an investigation built upon
weak conceptual foundations is unlikely to go anywhere, or to say anything
substantial. I have often found myself running in circles, multiplying the
number of measures performed, before finally leaving the keyboard alone to lay
down my thoughts on paper. On the other hand, when the thoughts have settled and
the question is clear, one can quickly make a substantial contribution.\\

A lesson painstakingly learned during this thesis is that \emph{thinking} the
city is as important as \emph{measuring} the city, or \emph{modeling} the city.
Concepts guide us and tell us what to measure, what to model. In the same way
measures and model can tell us what to think. But it would be very naive to
believe that scientific enquiries are fueled by the sole discussion between
measures and models. A quick inspection of the literature reveals, in fact, that
many studies are based upon an intuition, a pattern that the author has felt and
whose existence she is trying to prove on a quantitative basis. They are based
on an hypothesis, an intuition, a vision. Something difficult to explain with
words, but that I am sure the reader will recognise in his own research
practice.
