% Acknowledgements

\pdfbookmark[1]{Acknowledgements}{Acknowledgements} % Bookmark name visible in a PDF viewer


\begingroup

\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\chapter*{Acknowledgements} % Acknowledgements section text

\begin{flushright}{\slshape    
    I had the idea that the world's so full of pain \\
    it must sometimes make a kind of singing.} \\\medskip
    --- Robert Hass,\, \emph{Faint Music}~\cite{Hass:1998}
\end{flushright}


\bigskip


I would first and foremost like to express my gratitude to Alain Barrat, Michael
Batty, Pablo Jensen, Jean-Pierre Nadal and Lena Sanders for accepting to be in
my thesis committee. I can only imagine how daunting reading a PhD thesis must
be, and I am really grateful for your time and help. I sincerely hope this was
not as awful as reading a 200-pages technical document sounds.\\


This thesis would not have been the same without the supervision of Marc
Barthelemy. I still remember the sunny, winter morning of 2012 when the bus
first droppped me in the middle of nowhere, in the middle of fields, not far
from the Institut de Physique Th\'eorique. Little did I know then that I would
spend the next 3 years with a view not on fields or a forest, but on a wall.
In fact, I tried to convince myself that would only be temporary. This, combined with my aversion for long commutes,
goes a long way to tell how valuable I. I am grateful for how quickly you
trusted me and let me do my own things. I am grateful for your scientific
teachings, but also for teaching me the things there are no textbooks for, for
teaching me the ropes. If my publication list is so long, it is also because
Marc knew when to push me and how to push me. It was not always easy, but I
haven't regretted  my decision a single day. For all of this, I am forever
indebted to you.\\


Enfin---pardon my french---je tiens \`{a} remercier mes parents, qui, plus
peut-etre que les efforts combin\'es de toutes les personnes sus-cit\'ees, ont
amplement contribu\'e à cette th\`ese. Il m'arrive souvent de regarder en
arrière et de rejouer ma trajectoire personnelle si vous n'aviez pas \'et\'e
l\`a. Si vous n'aviez pas ete qui vous etes. Aucune ne se termine ici, devant
mon ordinateur, a tapper les derniers mots de cette th\`ese.  Récemment, un
collègue me demandait: ``Comment est-ce qu'un fils d'agriculteur fait pour
devenir physicien?''. Je n'en ai aucune idée, c'est juste arrivé. Mais une
grande partie de la réponse a à voir avec vous. Cet enfant, que vous avez trouvé
une nuit la t\^ete coll\'ee \`a la vitre, persuad\'e d'avoir d\'ecouvert une
deuxi\`eme lune, perplexe que l'on ne n'en lui ait jamais parl\'e---elle se
r\'ev\'elera \^etre une simple r\'eflexion dans la vitre. Cet enfant aurait pu
devenir n'importe qui. Je peux le dire maintenant, je suis fier de vous, et
infiniment reconnaissant. Une grande partie des compliments qui me sont faits
vous reviennent, et ils sont amplement mérités.\\


Little did I know 3 years ago that I would sit on the blue or red plastic on the
seats, reading articles . I will miss the RER B, and the two hours we spent
together every day. Your unpredictability, your characteristic... smell. Some of
my best ideas, and some of my most stressfull moments. I remember a few
important phone calls, and so many petty ones. A sort of love-hate relationship
started between us. I hate you for your unpredicability, your tendency to be
late when it's 3 degrees outside and the wind is blowing. Your tendency to be so
damn hot when it is already hot outside. Your robotic, feminine voice,
announcing an unquantifiable delay with the absolute voice of nothing. With no
affect, because it does not mean anything to you to have us trapped for a few
hours. Especially on Friday nights, when it is raining and we are soaked for
having waited for you through the 'Retard 10mins' sign. Because we had no
choice. But I forgive you, because I also had my best moments. I read; oh, I
read so much, craddled by the regular hops you make in the rusted tracks. I had
some good ideas with you too, as I was watching the banlieue passing by, rolling
away from me. I will miss you. Like one misses an abusive ex-partner, probably
more because of the routine you symbolised than your comfort and punctuality.\\

Finally, I would like to thank colleagues and friends. I know it was difficult
for you, too.\\

I wish to dedicate this thesis to the mistakes I made, to those I will
inevitably make. I can only hope they won't be the same.

%List of people to thank

%\begin{itemize}
	%\item Thomas. Un collègue en or, et surtout, un ami.
	%\item Hannah
	%\item Benoit
        %\item Hadrien
        %\item Cl\'ement
        %\item Bo
        %\item Oussama for being a real friend, with real values. A man if there
            %are any left on this planet.
        %\item K\'evin
    %\begin{quote}
        %Si on me presse de dire pourquoi je l'aimais, je sens que cela ne se peut
        %exprimer, qu'en r\'epondant : "Parce que c'\'etait lui, parce que c'\'etait moi."
    %\end{quote}
	%\item Riccardo and Giulia
	%\item secretaries
	%\item Ed
	%\item Arianna, Garvin, Elsa, Ed, Carlos
	%\item Marie and Olivia
	%\item Elisa
	%\item Bruno, Elisa, Sara, Giancarlo
	%\item Rosa-L\'y
        %\item The Linacre College Boat Club, for accepting me back in a boat,
            %despite my poor fitness after $3$ years of intense PhD. For letting
            %me row at Oxford's Summer VIII's and replace horrible thoughts about
            %my dissertation by equally horrible though about erging.
%\end{itemize}

%\noindent Many thanks to the Open Source community, without which I would not
%have been able to do even a half of this thesis. The developpers behind linux,
%for providing such a stable and efficient system. The developpers of Python and
%of the many python libraries. All the work I have done has been made possible by
%the selfless work of many before me.


\endgroup
